### 2016-09-06 FrontBackmatter/Introduction.tex
Write another section in the introduction

\chapter*{Introduction}
\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
\addcontentsline{toc}{chapter}{\tocEntry{Introduction}}

@@

\begin{epigraphs}
\qitem{
pity this busy monster, manunkind,
    }{
    --- \textsc{E. E. Cummings}}
\qitem{
[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
}{
--- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}}
\end{epigraphs}

@@

\todo[inline]{Quantum chemistry as they study of
molecular properties by means of \emph{ab initio} physico-mathematical
theories and computer programs.
Introduce the concept of model chemistries, but then mention that we are
interested in different aspects and will look at different "axes".}

This question was posed by R.~S.~Mulliken over a half-century ago, as
lecture title for the acceptance of the Gilbert N.~Lewis award in 1960.

@@

This will be the subject of Chapter \ref{ch:QM}

@@

Before introducing our model of choice, let us consider how an
environment might affect molecular observables of interest.
Environment effects are usually classified as:
\begin{itemize}
\item \emph{direct}.
  These effect stem straightforwardly from the modification underwent by
  the solute electronic density when interacting with the environment.
\item \emph{Indirect}.
  It is common for solutes to exhibit different minimum-energy
  conformations in different environments. These effects are commonly
  labelled as indirect.
\item \emph{Local field}.
  Light-matter interactions are also affected by the environment. Local
  modifications of externally applied fields subtly influence molecular
  responses.~\autocite{Cammi1998-jp, Pipolo2014-sd}
\item \emph{Dynamic}.
  Probing the nature of molecular excited states has enormous
  technological impact. The presence of the environment radically
  influences excited states, since relaxation processes in the medium
  become important.
\item \emph{Specific}. This catch-all category includes all effects
  stemming from the peculiar solute-solvent pair interactions that
  cannot be fully described under any of the previous labels.
  In general, modelling such effects demands an atomistic level of
  detail.
\end{itemize}

>>>>

\chapter*{Introduction}
\pdfbookmark[1]{Introduction}{introduction}

@@

\begin{epigraphs}
\setlength\epigraphwidth{7.5cm}
\qitem{
plays with the bigness of his littleness

--- electrons deify one razorblade

into a mountainrange; lenses extend

unwish through curving wherewhen till unwish

returns on its unself.
    }{
    --- \textsc{E. E. Cummings}}
\setlength\epigraphwidth{7cm}
\qitem{
[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
}{
--- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}}
\end{epigraphs}

@@

This question was posed by R.~S.~Mulliken over a half-century
ago\footnote{This was the title Mulliken chose for his Gilbert N.~Lewis
award acceptance speech in 1960.}
and can be considered the fundamental research question behind the
development of quantum chemistry.
The purpose of quantum chemistry is to provide models based on first
principles that can help \emph{understand} and \emph{predict} molecular properties.
As stated by \citeauthor{Dirac1929-gn} in his \citetitle{Dirac1929-gn}
paper:\autocite{Dirac1929-gn, Kutzelnigg2000-fl}
\blockquote{The underlying physical laws necessary for the mathematical
theory of a large part of physics and the whole of chemistry are thus
completely known, and the difficulty is only that the exact application of
these laws leads to equations much too complicated to be soluble. It
therefore becomes desirable that approximate practical methods of
applying quantum mechanics should be developed, which can lead to an
explanation of the main features of complex atomic systems without too
much computation.}
Following Dirac's \emph{dictum}, in quantum chemistry we apply physical
models based on quantum many-body methods to molecular systems, employ
their mathematical realizations and devise computable approximations.
The central idea is, in fact, to be able to obtain an algorithmic
implementation of the methods that can be applied to interesting
chemical systems: model building and software implementations are two
closely intertwined aspects in the practice of quantum chemistry.

@@

Quantum chemical methods have become invaluable tools in the modern
development of chemistry,\autocite{Lee1995-pw, Helgaker2004-oz,
Tajti2004-ye} as witnessed by the Nobel prizes awarded in 1998\autocite{Nobel1998}
and 2013.\autocite{Nobel2013}
The concept of \emph{model chemistries} is at the heart of these
successes. Introduced by \citeauthor{Pople1999-gt}, theoretical model
chemistries are specific combinations of approximations in the basis set
and molecular electronic structure methodology.\autocite{Pople1999-gt}
Model chemistries are \emph{systematically improvable} so that it is
possible to achieve the heaven of chemical accuracy by relaxing
approximations in the model, albeit at an increased computational cost.
Figure \ref{fig:RQC-axis} presents the usual depiction of this concept
as a set of orthogonal axes, where chemical accuracy can be achieved by
moving away from the origin.
Relativity can be considered as the third axis of quantum
chemistry: accuracy can be reached improving on the Hamiltonian, method
and basis axes.\autocite{Saue2011-qg}

The description of the ideas and methods at the basis of quantum
chemistry will be the subject of Chapter \ref{ch:QM}.
I will put emphasis on the methods that have been relevant in the work
presented in this thesis. Section \ref{sec:mqm} discusses some of the
ideas of quantum many-body theory relevant to quantum chemistry. Section
\ref{sec:mean-field} introduces the workhorses of quantum chemistry
\acrlong*{HF} and \acrlong*{KS} \acrlong*{DFT}.
The \acrlong*{MBPT} and \acrlong*{CC} rungs on the methodology axis will
be discussed in Sections \ref{sec:mbpt} and \ref{sec:coupled-cluster},
respectively.

@@

Before introducing our model of choice, let us consider how an
environment might affect molecular observables of interest.
Environment effects are usually classified as:
\begin{description}[font=\normalfont\scshape]
\item[Direct.]
  These effect stem straightforwardly from the modification underwent by
  the solute electronic density when interacting with the environment.
\item[Indirect.]
  It is common for solutes to exhibit different minimum-energy
  conformations in different environments. These effects are commonly
  labelled as indirect.
\item[Local field.]
  Light-matter interactions are also affected by the environment. Local
  modifications of externally applied fields subtly influence molecular
  responses.~\autocite{Cammi1998-jp, Pipolo2014-sd}
\item[Dynamic.]
  Probing the nature of molecular excited states has enormous
  technological impact. The presence of the environment radically
  influences excited states, since relaxation processes in the medium
  become important.
\item[Specific.] This catch-all category includes all effects
  stemming from the peculiar solute-solvent pair interactions that
  cannot be fully described under any of the previous labels.
  In general, modelling such effects demands an atomistic level of
  detail.
\end{description}
